---
title: "Plotting EA data"
author: "Brianna Hibner"
date: "2024-04-23"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
This file uploads data produced by the EA in the CUBES lab at CU Boulder (We acknowledge the analytical contributions of the CU Boulder Earth Systems Stable Isotope Lab (CUBES-SIL) Core Facility (RRID:SCR_019300)).


Data is filtered to just include samples from the ooid sand itself, ggplot boxplot function is used for both TOC and d13C. Further plot beautification occurred in adobe illustrator (position of legend, and combining plots)

#Loading packages
```{r packages}
library(ggplot2)
library(tidyr)
library(dplyr)
library(openxlsx)
library(LaCroixColoR)
```

## Opening data
```{r data}
#loading the xcel sheet (found in the 'data' folder)
Ooid_dat <- readxl::read_xlsx("data/Combined_clean_dat.xlsx", 
    #skipping first 3 line 
    skip=3 ) 
  # needed assistance for reading type of data
    col_types = c("text", "text","text","text","text", "numeric", "numeric", "numeric","numeric","numeric") 
    

```

```{r}
unrinsed_dat <- Ooid_dat |>
   #just ooid data
   filter(`TOC sample type`== 'ooid') |>
  #avoiding rinsed data
  filter(`extraction loca`== 'nut bag')
  
unrinsed_dat |>  
  filter(`Retrieval date`=='6 months')

```

##Plot of TOC data
```{r TOC plot}
#filtering for just the ooid sand data
Ooid_dat|>
  #just ooid data
   filter(`TOC sample type`== 'ooid') |>
  #avoiding rinsed data
  filter(`extraction loca`== 'nut bag') |>
  #making min and max error bars for TOC
  mutate(`%Cmin` = `%C` - 0.008)|>
  mutate(`%Cmax`= `%C` + 0.008) |>
  
#plotting
ggplot()+
  aes(x=`Retrieval date`, y=`%C`) + 
  
#error bars for TOC
  geom_errorbar(aes(ymin=`%Cmin`, ymax=`%Cmax`), color = 'lightgrey', width=0.20)+
  
    #boxplot base
  geom_boxplot()+
  
  #adding dots of sample color
  geom_point(aes(color = `bean bag sample`), size =2)+
  
 
  #color pallette 
   scale_color_manual(
    values=lacroix_palette("PassionFruit",
                           n = 6, 
                           type = "continuous"))+

      
  #reorganzing order of recovery date
  scale_x_discrete(limits = c("control", "1 week", "3 months", "6 months"))+
  
  #clean looking plot
  theme_light()


```


```{r d13C plots}
Ooid_dat|>
  #using just the ooid data 
  filter(`TOC sample type`== 'ooid') |>
    #avoiding rinsed data
  filter(`extraction loca`== 'nut bag') |>
    #making min and max error bars
  mutate(`d13Cmin` = `d13C` - 0.265)|>
  mutate(`d13Cmax`= `d13C` + 0.265) |>
  
#plotting
ggplot()+
  aes(x=`Retrieval date`, y=`d13C`) + 
  #error bars
  geom_errorbar(aes(ymin=`d13Cmin`, ymax=`d13Cmax`), color = 'lightgrey', width=0.20)+
  
  #boxplot base
  geom_boxplot()+
  #adding dots of sample color
  geom_point(aes(color = `bean bag sample`), size =2)+
    #color pallette
   scale_color_manual(
    values=lacroix_palette("PassionFruit",   n = 6, 
                           type = "continuous"))+

  
 #reorganzing order of recovery date
  scale_x_discrete(limits = c("control", "1 week", "3 months", "6 months"))+
  
  #gen theme
  theme_light() 

```


