---
title: "Raman data plots"
author: Lizzy Trower
format: html
embed-resources: true
editor: visual
theme: flatly
toc: true
---

## About

This code loads in and plots the Raman microspectroscopy data collected on sample CP13C.

## Load libraries

The LaCroixColoR package can be found at: <https://github.com/johannesbjork/LaCroixColoR>

Alternatively, the code can be modified to use another palette.

```{r load-libraries}
library(tidyverse)
library(LaCroixColoR)
```

## Load data

This code chunk loads the data from text files.

```{r load-data}
#load ROI-1 data
ROI1_01 <- read.table('CP13C_ROI1_1_2 s_785 nm_600 gr_mm_x50_01.txt')
ROI1_02 <- read.table('CP13C_ROI1_1_3 s_785 nm_600 gr_mm_x50_02.txt')
ROI1_03 <- read.table('CP13C_ROI1_1_3 s_785 nm_600 gr_mm_x50_03.txt')
ROI1_04 <- read.table('CP13C_ROI1_1_3 s_785 nm_600 gr_mm_x50_04.txt')
ROI1_05 <- read.table('CP13C_ROI1_1_3 s_785 nm_600 gr_mm_x50_05.txt')

#load ROI-2 data
ROI2_01 <- read.table('CP13C_ROI2_1_3 s_785 nm_600 gr_mm_x50_01.txt')
ROI2_02 <- read.table('CP13C_ROI2_1_3 s_785 nm_600 gr_mm_x50_02.txt')
ROI2_03 <- read.table('CP13C_ROI2_1_3 s_785 nm_600 gr_mm_x50_03.txt')
ROI2_04 <- read.table('CP13C_ROI2_1_3 s_785 nm_600 gr_mm_x50_04.txt')

#load ROI-3 data
ROI3_01 <- read.table('CP13C_ROI3_1_3 s_785 nm_600 gr_mm_x50_01.txt')
ROI3_02 <- read.table('CP13C_ROI3_2_3 s_785 nm_600 gr_mm_x50_02.txt')
ROI3_03 <- read.table('CP13C_ROI3_1_3 s_785 nm_600 gr_mm_x50_03.txt')
ROI3_04 <- read.table('CP13C_ROI3_1_3 s_785 nm_600 gr_mm_x50_04.txt')
ROI3_05 <- read.table('CP13C_ROI3_1_3 s_785 nm_600 gr_mm_x50_05.txt')

#load ROI-4 data
ROI4_01 <- read.table('CP13C_ROI4_1_3 s_785 nm_600 gr_mm_x50_01.txt')
ROI4_02 <- read.table('CP13C_ROI4_1_3 s_785 nm_600 gr_mm_x50_02.txt')

#load ROI-5 data
ROI5_01 <- read.table('CP13C_ROI5_2_3 s_785 nm_600 gr_mm_x50_01.txt')
```

## Data housekeeping

This code chunk adds some extra columns to keep track of which ROI and analysis each dataset corresponds to so they can be combined into a single dataframe for plotting.

```{r data-housekeeping}
ROI1_01$analysis <- rep(1,length(ROI1_01$V1))
ROI1_02$analysis <- rep(2,length(ROI1_02$V1))
ROI1_03$analysis <- rep(3,length(ROI1_03$V1))
ROI1_04$analysis <- rep(4,length(ROI1_04$V1))
ROI1_05$analysis <- rep(5,length(ROI1_05$V1))
ROI1_all <- rbind(ROI1_01,ROI1_02,ROI1_03,ROI1_04,ROI1_05)

ROI2_01$analysis <- rep(1,length(ROI2_01$V1))
ROI2_02$analysis <- rep(2,length(ROI2_02$V1))
ROI2_03$analysis <- rep(3,length(ROI2_03$V1))
ROI2_04$analysis <- rep(4,length(ROI2_04$V1))
ROI2_all <- rbind(ROI2_01,ROI2_02,ROI2_03,ROI2_04)

ROI3_01$analysis <- rep(1,length(ROI3_01$V1))
ROI3_02$analysis <- rep(2,length(ROI3_02$V1))
ROI3_03$analysis <- rep(3,length(ROI3_03$V1))
ROI3_04$analysis <- rep(4,length(ROI3_04$V1))
ROI3_05$analysis <- rep(5,length(ROI3_05$V1))
ROI3_all <- rbind(ROI3_01,ROI3_02,ROI3_03,ROI3_04,ROI3_05)

ROI4_01$analysis <- rep(1,length(ROI4_01$V1))
ROI4_02$analysis <- rep(2,length(ROI4_02$V1))
ROI4_all <- rbind(ROI4_01,ROI4_02)

ROI5_01$analysis <- rep(1,length(ROI5_01$V1))

ROI1_all$ROI <- rep(1,length(ROI1_all$V1))
ROI2_all$ROI <- rep(2,length(ROI2_all$V1))
ROI3_all$ROI <- rep(3,length(ROI3_all$V1))
ROI4_all$ROI <- rep(4,length(ROI4_all$V1))
ROI5_01$ROI <- rep(5,length(ROI5_01$V1))

alldata <- rbind(ROI1_all,ROI2_all,ROI3_all,ROI4_all,ROI5_01)
```

## Plot data

This code chunk plots all the data, with one subplot for each ROI. We add a dark gray dashed line to indicate the location of the aragonite L peak and a light gray dashed line to indicate the location of the calcite L peak.

```{r all-data-plot}
ROI.labs <- c('ROI 1','ROI 2','ROI 3','ROI 4','ROI 5')
names(ROI.labs) <- c('1','2','3','4','5')

ggplot(data = alldata, mapping = aes(x=V1,y=V2,color=factor(analysis))) +
  geom_vline(xintercept=206,color = 'darkgray',linetype = 'dashed') +
  geom_vline(xintercept=281,color = 'lightgray',linetype = 'dashed') +
  geom_line(show.legend = FALSE) + 
  facet_wrap(~ROI,
             labeller=labeller(ROI=ROI.labs)) +
  scale_color_manual(
    values=lacroix_palette(
      "PassionFruit", 
      type = "discrete")) +
  theme_light() +
  labs(y='Raman intensity',
       x=bquote('wavenumber (cm'^{-1}*')'))

ggsave('Ramanplot.pdf')
ggsave('Ramanplot.png')
```
